# Lab \| Data Visualization in R

## Load dataset and libraries

Load the Superstore dataset and libraries needed.

```{r}
super_store <- read.csv("dataset/Sample - Superstore.csv")
```

```{r}
#install.packages("VIM")
library(VIM)
library(ggplot2)
library(dplyr)
library(ggthemes)
```

## Boxplot

A summary table or boxplot can be useful to visualize the distribution of key variables. Generate a boxplot to visualize Sales and another to visualize Profit.

```{r}
#Boxplot for Sales
ggplot(super_store, aes(y=Sales)) +
  geom_boxplot() +
  labs(title= "Distribution of Sales", y= "Sales") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
# Boxplot for Profit
ggplot(super_store, aes(y = Profit)) +
  geom_boxplot() +
  labs(title = "Distribution of Profit", 
       y = "Profit") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Bar Plots

Create a bar plot to show the top 10 orders with highest value of sales.

```{r}
top_10_orders <- super_store %>%
  group_by(Order.ID) %>%
  summarise(Total_Sales= sum(Sales)) %>%
  arrange(desc(Total_Sales)) %>%
  slice_head(n=10)
print(top_10_orders)
```

```{r}
#Barplot Top 10 orders
ggplot(top_10_orders, aes(x = reorder(`Order.ID`, Total_Sales), y = Total_Sales)) +
  geom_bar(stat = "identity", fill = "maroon") +
  coord_flip() +  # Flip for better readability
  ggtitle("Top 10 Orders with Highest Sales") +
  xlab("Order ID") +
  ylab("Total Sales") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Heatmap

Use a heatmap to visualize the pattern of missing data.

``` r
# Install and load the VIM package if not already installed
library(VIM)

# Visualize missing data pattern
missing_pattern <- aggr(dataset, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(dataset), cex.axis=.7, gap=3, ylab=c("Missing data","Pattern"))
```

```{r}
missing_pattern <- aggr(super_store, 
                        col=c('navyblue', 'red'), 
                        numbers=TRUE, 
                        sortVars=TRUE, 
                        labels=names(super_store), 
                        cex.axis=0.7, 
                        gap=3, 
                        ylab=c("Missing data", "Pattern"))
```

## Histogram

Generate a histogram that can show the distribution of the Profit column.

```{r}
#Histogram Profit
ggplot(super_store, aes(x=Profit))+
  geom_histogram(binwidth =140, fill= 'green', color= 'black') +
  labs(title= "Distribution of Profit", x= "Profit" , y= "Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

## Bar Plots for Aggregated Data

Generate a bar plot that shows Total Sales by Category and a barplot that shows Profit by Category.

```{r}
#Total sales by category
sales_by_category <- super_store %>%
  group_by(Category) %>%
  summarise(
    Total_Sales=sum(Sales)
  )
ggplot(sales_by_category, aes(x = Category, y = Total_Sales, fill = Category)) +
  geom_bar(stat = "identity",color='black') +
  labs(title = "Total Sales by Category",
       x = "Category",
       y = "Total Sales") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +  
  scale_y_continuous(labels = scales::comma)+
  scale_fill_brewer(palette = "Set1")

```

```{r}
#Profit by category
profit_by_category <- super_store %>%
  group_by(Category) %>%
  summarise(
    Total_Profit= sum(Profit)
  )
ggplot(profit_by_category, aes(x = Category, y = Total_Profit, fill = Category)) +
  geom_bar(stat = "identity", color='black') +
  labs(title = "Total Profit by Category",
       x = "Category",
       y = "Total Profit") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +  
  scale_y_continuous(labels = scales::comma) +
  scale_fill_brewer(palette = "Set2")
```
